---
title: "Group 25 Assignment 7"
format: html
editor: visual
---

## Load Libraries and Data

First we load the data and the necessary libraries to perform a PCA

```{r}

library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)

biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

From the biopsy data set we can create a new data set that is appropriate to perform PCA analysis:

```{r}
PCA_dataset <- biopsy |> 
  select(-outcome) |> 
  prcomp(scale = TRUE)
```

The dataset below consists of all the features present in the biopsy dataset along with all the principal components

```{r}
Biopsy_Data_with_PCAcomponents <- PCA_dataset |> 
  augment(biopsy) 
```


## Data Analysis

### Variance explained

In this step we find the eigenvalues from the PCA, and use function tidy to convert this object into a tidy data frame. Each eigenvalue represents the variance explained by each principal component in the PCA.

```{r}
PCA_eigenvalues <- PCA_dataset |> 
  tidy(matrix = "eigenvalues")
```

By plotting the eigenvalues, we can visualize the variance explained.

```{r}
PCA_eigenvalues |>
  ggplot(aes(PC, percent)) + 
  geom_col(fill = "purple", alpha = 0.6) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.1))) +
  ggtitle("Variance explained by each principal component")
```

Based on the plot, the first principal component explains over 60 % of the variance.

#### First Principal Component

From the variance explained plot, we clearly see that the first pricinpial component is the most important. By plotting it:
```{r}
Biopsy_Data_with_PCAcomponents |> 
  ggplot(aes(x = .fittedPC1,
             color = outcome)) + 
  geom_density(size = 1.2,
               alpha = 0.5) +
  scale_color_manual(values = c("benign" = "blue",
                                "malignant" = "red")) +
  labs(x = "First Principal Component",
       color = "Outcome") +
  theme_minimal()
```

Although there is some overlap, the first PC can distinguish both outcomes pretty well.

#### First and Second Principal Components
Creating scatter plot of the two first fitted PCA columns:

```{r}
Biopsy_Data_with_PCAcomponents |>
  select(.fittedPC1, .fittedPC2, outcome) |>
  ggplot(aes(x = .fittedPC1,
             y = .fittedPC2,
             color = outcome)) +
  geom_point(alpha = 0.5) +
  labs(title = 'Scatter plot of PCA 1 and PCA 2',
       x = "First Principal Component",
       y = "Second Principal Component",
       color = "Outcome") +
  scale_color_manual( values = c(malignant = "red",
                                 benign = "darkblue")) +
  theme_minimal()
```
We see that with the first two principal components we can already distinguish pretty accurately both outcomes.


#### First and Third Principal Components

```{r}
Biopsy_Data_with_PCAcomponents |> 
  ggplot(aes(x = .fittedPC1,
             y = .fittedPC3,
             color = outcome)) +
  geom_point(alpha = 0.5) + 
  labs(title = 'Scatter plot of PCA 1 and PCA 3',
       x = "First Principal Component",
       y = "Third Principal Component") +
  scale_color_manual( values = c(malignant = "red",
                                 benign = "darkblue")) +
  theme_minimal()
  
```

We also see that the the 1st and 3rd PC differentiate pretty well a beningn outcome and a malignant outcome.


