---
title: "Group 25 Assignment 7"
format: html
editor: visual
---

## Background

## Load and Clean Data

First we load the data and the necessary libraries to perform a PCA

```{r}

library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)

biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

From the biopsy data set we can create a new data set that is appropriate to perform PCA analysis:

```{r}
PCA_dataset <- biopsy |> 
  select(-outcome) |> 
  prcomp(scale = TRUE)
```

The dataset below consists of all the features present in the biopsy dataset along with all the principal components

```{r}
Biopsy_Data_with_PCAcomponents <- PCA_dataset |> 
  augment(biopsy) 
  
```

## Data Analysis

## Variance explained

In this step we find the eigenvalues from the PCA, and use function tidy to convert this object into a tidy data frame. Each eigenvalue represents the variance explained by each principal component in the PCA.

```{r}
PCA_eigenvalues <- PCA_dataset %>%
  tidy(matrix = "eigenvalues")
```

By plotting the eigenvalues, we can visualize the variance explained.

```{r}
PCA_eigenvalues %>%
  ggplot(aes(PC, percent)) + 
  geom_col(fill = "purple", alpha = 0.6) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.1))) +
  ggtitle("Variance explained by each principal component")
```

Based on the plot, the first principal component explains over 60 % of the variance.

HEAD
We will now add a new binary variable where if the outcome is benign = 0 and if it is malignant = 1

```{r}
Biopsy_Data_with_PCAcomponents <- Biopsy_Data_with_PCAcomponents |> 
  mutate(binary_outcome = case_when(
    outcome == "benign"~0,
    outcome == "malignant"~1
  ))
```

```{r}
Biopsy_Data_with_PCAcomponents |> 
  ggplot(aes(.fittedPC1, binary_outcome )) + 
  geom_point() + 
  theme_minimal()
```

## Data Analysis

Creating scatter plot of the two first fitted PCA columns.

```{r}
Biopsy_Data_with_PCAcomponents |>
  select(.fittedPC1, .fittedPC2, outcome) |>  
  ggplot(aes(x = .fittedPC1, 
             y = .fittedPC2, 
             color = outcome)) +  
  geom_point() + 
  ggtitle('Scatter plot of PCA 1 and PCA 2') +
  scale_color_manual(
    values = c(malignant = "purple", benign = "darkblue"))  # Correct closing parenthesis

