---
title: "Group 25 Assignment 7"
format: html
editor: visual
---

## Background

## Load and Clean Data

First we load the data and the necessary libraries to perform a PCA

```{r}
#| warning: false
library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)

biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

From the biopsy data set we can create a new data set that is appropriate to perform PCA analysis:

```{r}
PCA_dataset <- biopsy |> 
  select(-outcome) |> 
  prcomp(scale = TRUE)
```

The dataset below consists of all the features present in the biopsy dataset along with all the principal components

```{r}
Biopsy_Data_with_PCAcomponents <- PCA_dataset |> 
  augment(biopsy) 
  
```

We will now add a new binary variable where if the outcome is benign = 0 and if it is malignant = 1

```{r}
Biopsy_Data_with_PCAcomponents <- Biopsy_Data_with_PCAcomponents |> 
  mutate(binary_outcome = case_when(
    outcome == "benign"~0,
    outcome == "malignant"~1
  ))
```

```{r}
Biopsy_Data_with_PCAcomponents |> 
  ggplot(aes(.fittedPC1, binary_outcome )) + 
  geom_point() + 
  theme_minimal()
```

## Data Analysis
