---
title: "Group 25 Assignment 7"
format: html
editor: visual
---

## Background

## Load and Clean Data

First we load the data and the necessary libraries to perform a PCA

```{r}
#| warning: false
library(tidyverse)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)

biopsy <- read_csv("https://wilkelab.org/classes/SDS348/data_sets/biopsy.csv")
```

From the biopsy data set we can create a new data set that is appropriate to perform PCA analysis:

```{r}
PCA_dataset <- biopsy |> 
  select(-outcome) |> 
  prcomp(scale = TRUE)
```

```{r}
Biopsy_Data_with_PCAcomponents <- PCA_dataset |> 
  augment(biopsy) 
  
```

## Data Analysis

## Variance explained

In this step we find the eigenvalues from the PCA, and use function tidy to convert this object into a tidy data frame. Each eigenvalue represents the variance explained by each principal component in the PCA.

```{r}
PCA_eigenvalues <- PCA_dataset %>%
  tidy(matrix = "eigenvalues")
```

By plotting the eigenvalues, we can visualize the variance explained.

```{r}
PCA_eigenvalues %>%
  ggplot(aes(PC, percent)) + 
  geom_col(fill = "purple", alpha = 0.6) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.1))) +
  ggtitle("Variance explained by each principal component")
```

Based on the plot, the first principal component explains over 60 % of the variance.
